---
title: "2018-2023-analysis"
output: html_document
---

```{r}
source("libraries.R")

data_path <- "K:/DataServices/Projects/Current_Projects/Regional_Plan_Update_Research/Speculative Investment/Data/"
setwd(data_path)

# homes for profit results with new llc network analysis methodology 2000-2023 data
hfp_networks_18_22 <- read_csv("20250109_warren_speculative-investment-analysis-dataset_withforeclosure_5yr-window-networks.csv") |> 
  filter(year >= 2018 & year <= 2022)

gc()
```


# Helper variables
```{r}
#Variables to help
res_list <- c("R1F", "R2F", "R3F", "CON")
#can't include first or last 3 years of data due to investor definitions
investor_year_min <- min(hfp_networks_18_22$year) + 4
investor_year_max <- max(hfp_networks_18_22$year) - 4
#can't include first 2 years or last 2 years due to flip definition
flip_year_min <- min(hfp_networks_18_22$year) + 2
flip_year_max <- max(hfp_networks_18_22$year) - 2
```


# Data points to compare to HFP report

## All 2018-2022
```{r}
count(hfp_networks_18_22)

#Percent of 2018-2022 transactions made by an investor
hfp_networks_18_22 |> 
  #calculate total transactions 
  mutate(total_transactions_period = n()) |>
  #filter to only investor transactions
  filter(investor_type_purchase != "Non-investor") |>
  #count total investor transactions then calculate %
  mutate(investor_transactions = n(),
         investor_transactions_p = investor_transactions/total_transactions_period) |>
  select(investor_transactions_p) |>
  distinct() 
```
- 514,079 transactions
- 21.2% investor transacations

## MAPC
```{r}
# count of transactions in MAPC region
count(filter(hfp_networks_18_22, mapc == 1))

#Percent of transactions in Metro Boston made by an investor - time period?
hfp_networks_18_22 |> 
  #filter to MAPC
  filter(mapc == 1) |>
  #calculate total transactions in investor years
  mutate(total_transactions_period = n()) |>
  #filter to only investor transactions
  filter(investor_type_purchase != "Non-investor") |>
  #count total investor transactions then calculate %
  mutate(investor_transactions = n(),
         investor_transactions_p = investor_transactions/total_transactions_period) |>
  select(investor_transactions_p) |>
  distinct() 
```
- 218,680 MAPC transactions
- 20.8% investor transactions

# Tables

## Helpful to know - count and % of investors by size and property type
```{r}
hfp_networks_18_22 |> 
  #filter to MAPC
  filter(mapc == 1) |>
  group_by(investor_type_purchase) |>
  mutate(trans_by_investor_type = n()) |>
  filter(restype %in% res_list | restype == 'APT') |>
  group_by(restype, investor_type_purchase) |>
  mutate(count = n(),
         percent = count/trans_by_investor_type) |>
  select(investor_type_purchase, restype, count, percent) |>
  arrange(investor_type_purchase, restype) |>
  distinct()
```


## Table 1: Warren Group Real Estate Transactions by Residential Type, MAPC Region
```{r}
hfp_networks_18_22 |> 
  #filter to MAPC
  filter(mapc == 1) |>
  mutate(total_transactions = n()) |> 
  select(restype, total_transactions) |>
  mutate(restype_group = ifelse(restype == "REO" | restype == "MOB" | restype == "MUR" | restype == "NEW" | restype == "OMR", "Other Residential Buildings", restype)) |> 
  group_by(restype_group) |>
  #count transactions by restype then divide by total transactions in full dataset
  reframe(
    transactions = n(),
    transactions_p = transactions/total_transactions
  ) |>
  distinct() |> 
  arrange(factor(restype_group, levels = c("R1F", "R2F", "R3F", "CON", "APT", "Other Residential Buildings")))
```


# Table 2: Warren Group Real Estate Investor Transactions by Investor Size, MAPC Region
```{r}
hfp_networks_18_22 |> 
  #filter to MAPC
  filter(mapc == 1) |>
  #filtering to only investor transactions
  filter(investor_type_purchase != "Non-investor") |>
  #count of all investor transatcionts
  mutate(total_investor = n()) |>
  group_by(investor_type_purchase) |>
  #count investor transactions by investor size and then calculate %
  reframe(transactions = n(),
          transactions_p = transactions/total_investor) |>
  distinct() |> 
    arrange(factor(investor_type_purchase, levels = c("Small", "Medium", "Large", "Institutional")))
```


```{r}

```

