---
title: "LLC Network Analysis"
author: "Alexa DeRosa"
date: "2024-10-25"
output: html_document
---

## Libraries and Data

```{r}
source("libraries.R")
#source("data-exploration.R")

data_path <- "K:/DataServices/Projects/Current_Projects/Regional_Plan_Update_Research/Speculative Investment/Data/"
#export_path <- "K:/DataServices/Projects/Current_Projects/Regional_Plan_Update_Research/Speculative Investment/"
setwd(data_path)

# homes for profit results under original methodology 2000-2023 data
hfp_original <- read_csv("20241220_warren_speculative-investment-analysis-dataset_withforeclosure_5yr-window.csv")
# homes for profit results with new llc network analysis methodology 2000-2023 data
hfp_networks <- read_csv("20250109_warren_speculative-investment-analysis-dataset_withforeclosure_5yr-window-networks.csv")

network_matches <- hfp_networks |> 
  filter(!is.na(network_group))

gc()
```

network-170 all medium

# add fields to help with analysis 
```{r}
hfp_networks <- hfp_networks |> 
  mutate(
    # add field to determine if an llc is part of a network
    network_flag = ifelse(is.na(network_group), 0, 1)
  )
```

# who are we matching to networks?
```{r}
count(hfp_networks, investor_type_purchase_llc, network_flag)

#llcs
hfp_networks |> 
  count(buyer_llc_ind, network_flag) |> 
  filter(buyer_llc_ind == 1) |> 
  mutate(
    total_t = sum(n),
    p_t = n/total_t
  )

#llps
hfp_networks |> 
  count(buyer_llp_ind, network_flag) |> 
  filter(buyer_llp_ind == 1) |> 
  mutate(
    total_t = sum(n),
    p_t = n/total_t
  )

# business
hfp_networks |> 
  count(buyer_bus_ind, network_flag) |> 
  filter(buyer_bus_ind == 1) |> 
  mutate(
    total_t = sum(n),
    p_t = n/total_t
  )

# trust
hfp_networks |> 
  count(buyer_trst_ind, network_flag) |> 
  filter(buyer_trst_ind == 1) |> 
  mutate(
    total_t = sum(n),
    p_t = n/total_t
  )

# gse
hfp_networks |> 
  count(buyer_gse_ind, network_flag) |> 
  filter(buyer_gse_ind == 1) |> 
  mutate(
    total_t = sum(n),
    p_t = n/total_t
  )

# gov
hfp_networks |> 
  count(buyer_gov_ind, network_flag) |> 
  filter(buyer_gov_ind == 1) |> 
  mutate(
    total_t = sum(n),
    p_t = n/total_t
  )

# banks
hfp_networks |> 
  count(buyer_bnk_ind, network_flag) |> 
  filter(buyer_bnk_ind == 1) |> 
  mutate(
    total_t = sum(n),
    p_t = n/total_t
  )

test <- count(hfp_networks, investor_type_purchase_llc, buyer_bus_ind, buyer_trst_ind, buyer_gse_ind, buyer_gov_ind, buyer_bnk_ind, network_flag)
```

Findings:
- we are matching 40,842 llc/llp transactions to networks
  - 39,989 llc transactions (42% of all LLC transactions)
  - 97 llp transactions (22% of all LLP transactions)
- we are matching 60,595 non-llc/llp entity transactions to networks
  - 6,724 business transactions (20% of all business transactions)
    - businesses are defined as buyers with 'INC', 'CORP', or 'CO' in the name
  - 8,864 transactions by trusts (11% of all trust transactions)
  - 15,866 gse transactions (72%)
  - 1 gov't transaction (.07%)
  - 37,122 bank transactions (87% of all bank transactions)
  - there are 1,330 transactions being pulled into networks that don't meet any of the above flags
  
```{r}
hfp_networks |> 
  filter(investor_type_purchase_llc != "Small LLC" & buyer_trst_ind == 0 & buyer_bus_ind == 0 & buyer_gse_ind == 0 &
           buyer_gov_ind == 0 & buyer_bnk_ind == 0 & network_flag == 1) |> 
  select(buyer1_adj, address, restype, municipal, investor_type_purchase) |> 
  count(buyer1_adj) |> 
  arrange(desc(n))
```
The 1,330 transactions are from 413 different buyers. The top 2 buyers have significantly more transactions than the other 411, FREMONT INVESTMENT AND LOAN (134), FIRST HORIZON HOME LOANS (118). These 413 buyers appear to be mostly mortgage companies, investment groups, colleges, and housing authorities with one banks not being caught by the bank index (because we aren't searching for 'BANK '), lots of buyers now look like "CONDOMINIUM JOHN" or "CONDOMINIUM WILLIAM F"

# how many llc/llp transactions are we matching per year
```{r}
annual_counts <- hfp_networks |> 
  # filter to just llc/llp transactions
  filter(buyer_llc_ind == 1 | buyer_llp_ind == 1) |> 
  count(year) |> 
  rename(total_llc = n)

annual_counts_matched <- hfp_networks |> 
  # filter to just llc/llp transactions that are in networks
  filter((buyer_llc_ind == 1 | buyer_llp_ind == 1) & network_flag == 1) |> 
  count(year) |> 
  rename(matched_llc = n)

counts <- full_join(annual_counts, annual_counts_matched) |> 
  mutate(dif = total_llc - matched_llc,
         percent = matched_llc/total_llc)

ggplot(counts, aes(x = year, y = percent*100)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label= paste0((round(percent*100, digits = 0)), "%"), vjust = -.5), size = 3)
```

Findings: 
- we're matching the highest percentage of transactions in 2021 - 62% (4,950 llc transactions matched)
- we're only matching 44% of llc transactions in 2023, likely because parcel data has not been updated in many towns
- should we go back to 2015 (43%) or 2016 (46%) for this analysis?

# how many unique llcs are we matching 
```{r}
# total
total_llcs <- hfp_networks |> 
  # filter to just llc/llp transactions
  filter(buyer_llc_ind == 1 | buyer_llp_ind == 1) |>
  select(buyer1_adj) |> 
  distinct()
  # 54,201

matched_llcs <- hfp_networks |> 
  # filter to just llc/llp transactions that are in networks
  filter((buyer_llc_ind == 1 | buyer_llp_ind == 1) & network_flag == 1) |> 
  select(buyer1_adj) |> 
  distinct()
# 17,834

# total entities
total_ent <- hfp_networks |> 
  # filter to just non llc/llp entity transactions
  filter(buyer_bus_ind == 1 | buyer_trst_ind == 1 | buyer_gse_ind == 1 | buyer_gov_ind == 1 | buyer_bnk_ind == 1) |>
  select(buyer1_adj) |> 
  distinct()
  # 66,407

matched_ent <- hfp_networks |> 
  # filter to just non llc/llp entity transactions that are in networks
  filter((buyer_bus_ind == 1 | buyer_trst_ind == 1 | buyer_gse_ind == 1 | buyer_gov_ind == 1 | buyer_bnk_ind == 1) 
         & network_flag == 1) |> 
  select(buyer1_adj, network_group) |> 
  distinct()
# 1,466

# 2018-2022
total_llcs21 <- hfp_networks |> 
  filter((buyer_llc_ind == 1 | buyer_llp_ind == 1) & (year >= 2018 & year <= 2022)) |> 
  select(buyer1_adj) |> 
  distinct()
# 2021: 6,003 2018-2022: 21,500

matched_llcs21 <- hfp_networks |> 
  filter((buyer_llc_ind == 1 | buyer_llp_ind == 1)  & (year >= 2018 & year <= 2022) & network_flag == 1) |> 
  select(buyer1_adj) |> 
  distinct()
#2021: 3,504 2018-2022: 11,017 (51)
```
Findings:
matching 58% of unique LLCs in 2021, 33% of LLCs overall

# how do investors in networks compare to those that aren't in networks? (2021) -network id
```{r}
non_matches21h <- hfp_networks |> 
  filter(network_flag == 0 & (year >= 2018 & year <= 2022)) |> 
  count(investor_type_purchase) |> 
  rename(no_network_count = n) 

network_matches21h <- hfp_networks |> 
  filter(network_flag == 1 & (year >= 2018 & year <= 2022)) |> 
  count(investor_type_purchase) |> 
  rename(network_count = n)

inv_typ_21h <- full_join(non_matches21h, network_matches21h, by = "investor_type_purchase") |> 
  filter(investor_type_purchase != "Non-investor")

inv_typ_21h |> 
  pivot_longer(
    !investor_type_purchase,
    names_to = "network_type",
    values_to = "count"
  ) |> 
  ggplot(aes(x = investor_type_purchase, y = count, fill = network_type)) + 
  geom_bar(position = "dodge", stat = "identity", color = "black")
```

########################### Comparing the old and new methodologies ######################

joining data output for original methodology with the new methodology
```{r}
hfp_networks_select <- hfp_networks |> 
  select(id, buyer1_adj, final_name, investor_type_purchase_count, investor_type_purchase_llc, investor_type_purchase_building, investor_type_purchase_value, investor_type_purchase, network_flag) |> 
  # need to rename columns before join
  rename_with(.fn = ~ paste0("n_", .x)) |> 
  rename(id = n_id)

hfp_joined <- left_join(hfp_original, hfp_networks_select, by = "id") 

rm(hfp_networks_select)
gc()

```

# how do investors in networks compare to those that aren't in networks? (2021) -buyer1_adj
```{r}
o_non_matches21h <- hfp_joined |> 
  filter(n_network_flag == 0 & (year >= 2018 & year <= 2022)) |> 
  count(investor_type_purchase) |> 
  rename(no_network_count = n) 

o_network_matches21h <- hfp_joined |> 
  filter(n_network_flag == 1 & (year >= 2018 & year <= 2022)) |> 
  count(investor_type_purchase) |> 
  rename(network_count = n)

o_inv_typ_21h <- full_join(o_non_matches21h, o_network_matches21h, by = "investor_type_purchase") |> 
  filter(investor_type_purchase != "Non-investor")

o_inv_typ_21h |> 
  pivot_longer(
    !investor_type_purchase,
    names_to = "network_type",
    values_to = "count"
  ) |> 
  ggplot(aes(x = investor_type_purchase, y = count, fill = network_type)) + 
  geom_bar(position = "dodge", stat = "identity", color = "black")
```

## How did investors move sizes
```{r}
hfp_comp <- hfp_joined |> 
  # select(id, investor_type_purchase_count, investor_type_purchase_llc, investor_type_purchase_building, investor_type_purchase_value, investor_type_purchase, n_investor_type_purchase_count, n_investor_type_purchase_llc, n_investor_type_purchase_building, n_investor_type_purchase_value, n_investor_type_purchase) |> 
  mutate(
    #make columns with new and original investor types
    inv_type_pur_comp = paste0(investor_type_purchase, " to ", n_investor_type_purchase), 
    inv_type_pur_llc_comp = paste0(investor_type_purchase_llc, " to ", n_investor_type_purchase_llc),
    inv_type_pur_bldg_comp = paste0(investor_type_purchase_building, " to ", n_investor_type_purchase_building),
    inv_type_pur_count_comp = paste0(investor_type_purchase_count, " to ", n_investor_type_purchase_count),
    inv_type_pur_value_comp = paste0(investor_type_purchase_value, " to ", n_investor_type_purchase_value),
    size_change = ifelse(inv_type_pur_comp == "Institutional to Institutional" | inv_type_pur_comp == "Large to Large" | 
                           inv_type_pur_comp == "Medium to Medium" | inv_type_pur_comp == "Small to Small" | 
                           inv_type_pur_comp == "Non-investor to Non-investor", 0, 1)
  ) |> 
  select(id, year, buyer1_adj, n_buyer1_adj, final_name, investor_type_purchase_llc, inv_type_pur_comp, inv_type_pur_llc_comp,
         inv_type_pur_bldg_comp, inv_type_pur_count_comp, inv_type_pur_value_comp, size_change)

#rm(hfp_joined)
gc()
## All buyers
#how many changes total
count(hfp_comp, size_change)
#how many changes 2015 on
hfp_comp |> filter(year >= 2015) |> count(size_change)
#how many changes for years where we match > 50% of LLCs
hfp_comp |> filter(year >= 2018 & year <= 2022) |> count(size_change)

## LLC buyers
#how many changes total
hfp_comp |> filter(investor_type_purchase_llc == "Small LLC") |> count(size_change)
#how many changes 2015 on
hfp_comp |> filter(investor_type_purchase_llc == "Small LLC" & year >= 2015) |> count(size_change)
#how many changes for years where we match > 50% of LLCs
hfp_comp |> filter(investor_type_purchase_llc == "Small LLC" & year >= 2018 & year <= 2022) |> count(size_change)
```
Findings:
- There were 17,031 transactions in the entire dataframe that saw a change in their final buy-side investor status (0.7% of all transactions)
  - 10,130 transactions from 2015-2023 (1.1% of transactions)
  - 6,623 transactions 2018-2022 (years where we match > 50% of LLCS) (1.3%)
- There were 12,473 LLC/LLP transactions in the entire data frame that saw a change in their final buy-side investor status (13.2% of all LLC transactions)
  - 8,954 transactions from 2015-2023 (15.5%)
  - 6,015 transactions 2018-2022 (21.5%)

## what do those changes look like  

# Overall investors
```{r}
#transactions
hfp_comp |> count(inv_type_pur_comp) |> arrange()

hfp_comp |> filter(year>=2018 & year<= 2022) |> count(inv_type_pur_comp) |> arrange()

#unique investors
hfp_comp |> select(-c(id, year, buyer1_adj)) |> distinct() |> count(inv_type_pur_comp) |> arrange()
```
Findings: overall investors
- no investors moved down a size!
- transaction level
  - 2,004 non-investor transactions moved up into investor categories!
  - 9,255 small investor transactions moved up to larger investor size, more than half became institutional
  - 2,954 medium investor transactions moved up to a larger size
  - 2,818 large investor transactions became institutional investor transactions
- unique investor level
  - 1,538 non-investors became investors, most became small investors
  - 7,850 investors moved up to larger sizes
    - 6,638 small investors moved up to larger sizes - more than half became institutional
    - 752 medium investors moved up
    - 460 large investors became institutional

# LLC investors
```{r}
hfp_comp |> count(inv_type_pur_llc_comp)

new_llcs <- hfp_comp |> filter(inv_type_pur_llc_comp == "Non-Small LLC to Small LLC")
```
Findings: LLC investors
- 3 transactions moved from being a Small LLC to Non-Small LLC - look into this
- 2,152 transactions went from Non-Small LLC to Small LLC - this is also weird really there shouldn't have been movement in this category
  
# Building investors
```{r}
hfp_comp |> count(inv_type_pur_bldg_comp)
```
Findings: Building Investors
- no transactions moved from building investor to non-investor!! 
- 10,051 transactions moved from non-building investors to building investors once in networks

# Count investors
```{r}
#transactions
hfp_comp |> count(inv_type_pur_count_comp) |> arrange()

#unique investors
hfp_comp |> select(-c(id, year, buyer1_adj)) |> distinct() |> count(inv_type_pur_count_comp) |> arrange()
```
Findings: Count investors
- no investors moved down a size!
- transaction level
  - 9,971 non-investor transactions moved up into count investor categories!
  - 2,168 small investor transactions moved up to larger count investor size, about half became institutional
  - 2,627 medium investor transactions moved up to a larger size
  - 2,503 large investor transactions became institutional investor transactions
- unique investor level
  - 7,929 non-investors became investors, a fairly even split into each of the four sizes !!!!
  - 1,451 investors moved up to larger sizes
    - 640 small investors moved up to larger sizes - more than half became institutional
    - 515 medium investors moved up
    - 296 large investors became institutional
    
# Value investors
```{r}
#transactions
hfp_comp |> count(inv_type_pur_value_comp) |> arrange()

#unique investors
hfp_comp |> select(-c(id, year, buyer1_adj)) |> distinct() |> count(inv_type_pur_value_comp) |> arrange()
```
Findings: Value investors
- no investors moved down a size!
- transaction level
  - 12,046 non-investor transactions moved up into count investor categories!
  - 2,551 small investor transactions moved up to larger count investor size, about half became institutional
  - 2,741 medium investor transactions moved up to a larger size
  - 2,618 large investor transactions became institutional investor transactions
- unique investor level
  - 6,703 non-investors became investors, more to large and institutional than medium and small
  - 1,100 investors moved up to larger sizes
    - 509 small investors moved up to larger sizes - more than half became institutional
    - 400 medium investors moved up
    - 191 large investors became institutional   

# how many investors and investor transactions did we have previously?
```{r}
#transactions
hfp_original |> filter(investor_type_purchase != "Non-investor") |> count()
hfp_networks |> filter(investor_type_purchase != "Non-investor") |> count()

#unique buyers
hfp_original |> select(final_name) |> distinct() |> count()
hfp_networks |> select(final_name) |> distinct() |> count()

#unique investors
hfp_original |> select(final_name, investor_type_purchase) |> distinct() |> filter(investor_type_purchase != "Non-investor") |> count()
hfp_networks |> select(final_name, investor_type_purchase) |> distinct() |> filter(investor_type_purchase != "Non-investor") |> count()

# difference between buyer1_adj count and final_name
hfp_original |> select(buyer1_adj) |> distinct() |> count() #1,696,313 original 
x <- hfp_networks |> select(buyer1_adj) |> distinct() |> count()#1,693,967 networks buyer1
y <- hfp_networks |> select(final_name) |> distinct() |> count() #1,686,690 networks final_name
x-y # 7,277

#what does this join look like?
buyers <- hfp_joined |> select(id, buyer1_adj, n_buyer1_adj, n_final_name) |> filter(buyer1_adj != n_buyer1_adj) |> 
  select(-c(id)) |> distinct()

```
- originally had 500,286 investor purchases of 2,351,320 (21.3%), now we have 502,290 (21.4%) (+2004 transactions)
- originally had 145,236 unique investors of 1,696,313 unique buyers (8.5% of buyers)
- now we have 137,593 unique investors of 1,686,690 unique buyers (8.2%)
    - 9,623 fewer unique investors with networks
    ?? why are there more unique buyers with the networks than there were before?
      - it looks like some of the name cleaning helped match buyer names, e.g. network-9972 FAGGIO LP and FAGGIO, L P were grouped together as FAGGIO LLP using Eric's methods