---
title: "Template Project"
author: "Zoe Iacovino and Alexa DeRosa"
date: "2024-10-25"
output: html_document
---

## Libraries and Functions

```{r}
source("libraries.R")
source("data-exploration.R")

network_matches <- warren_networks |> 
  filter(!is.na(network_group))
```

warren_networks is all transactions by llcs and llps from 2000-2023

# how many llcs transactions are we matching per year
```{r}
annual_counts <- count(warren_networks, year) |> 
  rename(total_llc = n)

annual_counts_matched <- count(network_matches, year) |> 
  rename(matched_llc = n)

counts <- full_join(annual_counts, annual_counts_matched) |> 
  mutate(dif = total_llc - matched_llc,
         percent = matched_llc/total_llc)

ggplot(counts, aes(x = year, y = percent)) +
  geom_bar(stat = "identity")
```

we're matching the highest percentage of transactions in 2021 - 62% (2,831 transactions not being matched)

# how many unique llcs are we matching 
```{r}
# total
total_llcs <- warren_networks |> 
  select(buyer1_adj) |> 
  distinct()
# 51,383

matched_llcs <- network_matches |> 
  select(buyer1_adj) |> 
  distinct()
# 17,243

# 2021
total_llcs21 <- warren_networks |> 
  filter(year == 2021) |> 
  select(buyer1_adj) |> 
  distinct()
# 5,624

matched_llcs21 <- network_matches |> 
  filter(year == 2021) |> 
  select(buyer1_adj) |> 
  distinct()
# 3,317
```

only matching 59% of unique llcs in 2021


# what do we know about the llcs we're matching

2021 data exploration (transactions)
```{r}
non_matches21 <- warren_networks |> 
  filter(is.na(network_group) & year == 2021) |> 
  count(investor_type_purchase) |> 
  rename(no_network_count = n) 

network_matches21 <- network_matches |> 
  filter(year == 2021) |> 
  count(investor_type_purchase) |> 
  rename(network_count = n)

inv_typ_21 <- full_join(non_matches21, network_matches21, by = "investor_type_purchase") |> 
  filter(investor_type_purchase != "Non-investor")

x <- inv_typ_21 |> 
  pivot_longer(
    !investor_type_purchase,
    names_to = "network_type",
    values_to = "count"
  ) |> 
  ggplot(aes(x = investor_type_purchase, y = count, fill = network_type)) + 
  geom_bar(position = "dodge", stat = "identity", color = "black")

x
```


# how do the networks change the investor sizes?
```{r}
hfp_networks <- read_csv("K:/DataServices/Projects/Current_Projects/Regional_Plan_Update_Research/Speculative Investment/Data/20241025_warren_speculative-investment-analysis-dataset_withforeclosure_5yr-window-networks.csv")

non_matches21h <- hfp_networks |> 
  filter(is.na(network_group) & year == 2021) |> 
  count(investor_type_purchase) |> 
  rename(no_network_count = n) 

network_matches21h <- hfp_networks |> 
  filter(!is.na(network_group) & year == 2021) |> 
  count(investor_type_purchase) |> 
  rename(network_count = n)

inv_typ_21h <- full_join(non_matches21, network_matches21, by = "investor_type_purchase") |> 
  filter(investor_type_purchase != "Non-investor")

y <- inv_typ_21h |> 
  pivot_longer(
    !investor_type_purchase,
    names_to = "network_type",
    values_to = "count"
  ) |> 
  ggplot(aes(x = investor_type_purchase, y = count, fill = network_type)) + 
  geom_bar(position = "dodge", stat = "identity", color = "black")

y
```





## Defining Variables

```{r}

# output paths
output_path<- paste0("C:/Users/",Sys.info()["username"],"/OneDrive - Metropolitan Area Planning Council")

arc_pro_gdb<- "K:/DataServices/Projects/Current_Projects/"

# geographics study area
study_area<- arc.open("K:/DataServices/Projects/Current_Projects/")

study_area<-arc.data2sf(arc.select(study_area))

# municipal filtering variables
muni_lower<- "Milford"

muni_upper<- "MILFORD"

muni_id_ob<- 185


```

