---
title: "annual-agg"
output: html_document
---
script for below Brandeis dat request

- MAPC communities, aggregated annually, at the tract level: If possible, both 2010 and 2020 tracts. If that's not possible to do both, then 2010 only.
- median sale price at the tract level annually
- all investors / investor non-flips / all flips / investor flips
- primary variables of interest: investor purchases, flips, and cash, so if there aren't enough observations for the other variables, like small/med/large/instit investors, then it would be fine to omit them.


```{r, warning=FALSE}
library(tidyverse)
options('scipen' = 10)

data_path = "K:/DataServices/Projects/Current_Projects/Regional_Plan_Update_Research/Speculative Investment/Data/"

setwd(data_path)

warren <- read_csv("2000_2025_warren_speculative-investment-analysis-dataset_withforeclosure_5yr-window-networks.csv")
```

 
```{r}
warren_mapc <- warren |> filter(!is.na(mapc_submarket))

warren_select <- warren_mapc |>  
  select(ID, muni_id, municipal, ct20_id, year, restype, cash_sale, price, price_adj, flip_ind, investor_type_purchase_count, investor_type_sale_count, investor_type_purchase_llc, investor_type_sale_llc, investor_type_purchase_building, investor_type_sale_building, investor_type_purchase_value, investor_type_sale_value, investor_type_purchase, investor_type_sale) %>% 
  mutate(municipal = case_when(
    municipal == 'Attleborough' ~ 'Attleboro',
    municipal == 'Manchester-by-the-Sea' ~ 'Manchester',
    TRUE ~ municipal
  ))

rm(warren)
gc()
```

```{r, warning=FALSE}
create_tables <- function(df, geog){
  warren_all = df %>% 
    select(muni_id,{{geog}}, year) %>% 
    distinct()
  
  transactions_all = df %>% 
    group_by(!!rlang::ensym(geog), year) %>% 
    summarize(sales_frq = n())
  
  trans_filtered = df |> 
    mutate(sales2_id = case_when(restype == "CON" & price >= 50000 & price <= 15000000 ~ 1,
                                  restype != "CON" & price >= 50000 & price <= 100000000 ~ 1)) |> 
    filter(sales2_id == 1) |> 
    group_by(!!rlang::ensym(geog), year) %>% 
    summarize(sales_frq2 = n(),
              med_pr = sum(price),
              med_pr_adj = sum(price_adj)) |> 
    mutate(med_pr = ifelse(sales_frq2 < 5, NA, med_pr),
           med_pr_adj = ifelse(sales_frq2 < 5, NA, med_pr_adj))
  
  #creating needed variables
  data_summary = df |> 
    mutate(
      inv_trans_f = ifelse(investor_type_purchase != 'Non-investor', 1, 0),
      sm_inv_f = ifelse((year >= 2004 & year <= 2021) & investor_type_purchase == 'Small', 1, 0),
      med_inv_f = ifelse((year >= 2004 & year <= 2021) & investor_type_purchase == 'Medium', 1, 0),
      lrg_inv_f = ifelse((year >= 2004 & year <= 2021) & investor_type_purchase == 'Large', 1, 0),
      inst_inv_f = ifelse((year >= 2004 & year <= 2021) & investor_type_purchase == 'Institutional', 1, 0),
      inv_llc_f = ifelse(investor_type_purchase_llc == 'Small LLC', 1, 0),
      v_inv_f = ifelse(investor_type_purchase_value != 'Non-value investor', 1, 0),
      c_inv_f = ifelse((year >= 2004 & year <= 2021) & investor_type_purchase_count != 'Non-count investor', 1, 0),
      b_inv_f = ifelse(investor_type_purchase_count != 'Non-count investor', 1, 0),
      flip_f = ifelse((year >= 2002 & year <= 2021) & flip_ind == 1, 1, 0),
      con_f = ifelse((year >= 2004 & year <= 2021) & restype == 'CON'),
      con_inv_f = ifelse((year >= 2004 & year <= 2021) & restype == 'CON' & investor_type_purchase != 'Non-investor'),
      sf_f = ifelse((year >= 2004 & year <= 2021) & restype == 'R1F'),
      sf_inv_f = ifelse((year >= 2004 & year <= 2021) & restype == 'R1F' & investor_type_purchase != 'Non-investor'),
      r2f_f = ifelse((year >= 2004 & year <= 2021) & restype == 'R2F'),
      r2f_inv_f = ifelse((year >= 2004 & year <= 2021) & restype == 'R2F' & investor_type_purchase != 'Non-investor'),
      r3f_f = ifelse((year >= 2004 & year <= 2021) & restype == 'R3F'),
      r3f_inv_f = ifelse((year >= 2004 & year <= 2021) & restype == 'R3F' & investor_type_purchase != 'Non-investor'),
      inv_non_flip_f = ifelse((year >= 2002 & year <= 2021) & investor_type_purchase != 'Non-investor' & flip_ind == 0, 1, 0),
      non_inv_flip_f = ifelse((year >= 2002 & year <= 2021) & flip_ind == 1 & investor_type_purchase == 'Non-investor')
    ) |> 
    group_by(!!rlang::ensym(geog), year) |>  
    summarize(inv_trans = sum(inv_trans_f),
              sm_inv = sum(sm_inv_f),
              med_inv = sum(med_inv_f),
              lrg_inv = sum(lrg_inv_f),
              inst_inv = sum(inst_inv_f),
              llc_inc = sum(inv_llc_f),
              value_inv = sum(v_inv_f),
              count_inv = sum(c_inv_f),
              build_inv = sum(b_inv_f),
              cash_trans = sum(cash_sale),
              flip_count = sum(flip_f),
              inv_non_flip = sum(inv_non_flip_f),
              non_inv_flip = sum(non_inv_flip_f),
              con_trans = sum(con_f),
              con_inv = sum(con_inv_f),
              sf_trans = sum(sf_f),
              sf_inv = sum(sf_inv_f),
              r2f_trans = sum(r2f_f),
              r2f_inv = sum(r2f_inv_f),
              r3f_trans = sum(r3f_f),
              r3f_inv = sum(r3f_inv_f),
              )


#join summary tables
warren_join = full_join(warren_all, transactions_all, by = c({{geog}}, 'year')) |> 
    full_join(trans_filtered, by = c({{geog}}, 'year')) |> 
    full_join(data_summary, by = c({{geog}}, 'year')) |> 
    #calculating %s
    mutate(inv_p = inv_trans/sales_frq, #inv trans/tot trans
           sm_inv_p = sm_inv/inv_trans, #small inv trans/tot inv trans
           med_inv_p = med_inv/inv_trans, #med inv trans/tot inv trans
           lrg_inv_p = lrg_inv/inv_trans, #large inv trans/tot inv trans
           inst_inv_p = inst_inv/inv_trans, #inst inv trans/tot inv trans
           li_inv = lrg_inv + inst_inv, #tot large and inst trans
           li_inv_p = li_inv/inv_trans, #lrg + inst inv trans/tot inv trans
           llc_inv_p = llc_inv/inv_trans, #llc inv trans/tot inv trans
           val_inv_p = value_inv/inv_trans, #value inv trans/tot inv trans
           c_inv_p = count_inv/inv_trans, #count investor transactions/tot inv trans
           bld_inv_p = build_inv/inv_trans, #building inv trans/tot inv trans
           cash_p = cash_trans/sales_frq, #cash trans / tot trans
           flip_p = flip_count/sales_frq, #flip trans / tot trans
           con_inv_p = con_inv/con_trans, 
           sf_inv_p = sf_inv/sf_trans,
           r2f_inv_p = r2f_inv/r2f_trans,
           r3f_inv_p = r3f_inv/r3f_trans,
    ) %>%
      relocate("li_inv", .after = inst_inv) %>%
      relocate(c("flip_count", "con_inv", "sf_inv", "r2f_inv", "r3f_inv"), .after = cash_trans) %>%
      relocate(c("flip_p", "con_inv_p", "sf_inv_p", "r2f_inv_p", "r3f_inv_p"), .after = cash_p) |> 
      #overwriting data for geographies with less then 5 sales
      mutate(
        med_pr = ifelse(sales_frq2 < 5, NA, med_pr),
        med_pr_adj = ifelse(sales_frq2 < 5, NA, med_pr)
      )

return(warren_join)
}

```


```{r, message=FALSE, warning=FALSE}
setwd(data_path)

#create tables
#run municipal tables
annual_muni_agg <- create_tables(warren_select, "municipal")

muni_publish <- annual_muni_agg |>
  arrange(municipal, year)

#export muni table
write.csv(muni_publish, 'warren_municipal_annual_agg_2000_2025.csv')
```

```{r, message=FALSE}
setwd(data_path)

#run ct tables
annual_ct_agg <- create_tables(warren_select, 'ct20_id')

ct_publish <- annual_ct_agg |> 
   arrange(ct20_id, year) |> 
   select(-c(muni_id))

#export ct table
write.csv(ct_publish, 'warren_ct_annual_agg_2000_2025.csv')
```

