---
title: "annual-agg"
output: html_document
---
script for below Brandeis data request

Notes on results: 
- there are 385 2020 census tract + year combinations across 63 census tracts where median sales price is suppressed due to there being fewer than 5 sales in that census tract in the given year
- there are 254 2010 census tract + year combinations across 43 census tracts where median sales price is suppressed due to there being fewer than 5 sales in that census tract in the given year


Potential issue: There are some count fields that are NA, causing some percent fields (denoted by _p) to be NaN. It seems like there are some transactions missing an individual investor classification field which also seems to cause an NA in the overall investor classification. This seems to be an issue going back to the original Homes for Profit scripts and not with this code. This issue impacts about 4% of transaction records within the MAPC region.

```{r, warning=FALSE}
library(tidyverse)
options('scipen' = 10)

data_path = "K:/DataServices/Projects/Current_Projects/Regional_Plan_Update_Research/Speculative Investment/Data/"
export_path = "K:/DataServices/Projects/Data_Requests/2026/Brandeis_HFP"

setwd(data_path)

#starting from mapc filtered file
warren_mapc <- read_csv("2000_2025_warren_speculative-investment-analysis-dataset_mapc_withforeclosure_5yr-window-networks.csv")
```

 
```{r}
#select only needed variables
warren_select <- warren_mapc |>  
  select(ID, ct20_id, ct10_id, year, restype, cash_sale, price, price_adj, flip_ind, investor_type_purchase_count, investor_type_sale_count, investor_type_purchase_llc, investor_type_sale_llc, investor_type_purchase_building, investor_type_sale_building, investor_type_purchase_value, investor_type_sale_value, investor_type_purchase, investor_type_sale)

rm(warren_mapc)
gc()
```

```{r, warning=FALSE}
create_tables <- function(df, geog){
  #table to save all munis/geographies/years
  warren_all = df %>% 
    select({{geog}}, year) %>% 
    distinct()
 
   #total transaction count
  transactions_all = df %>% 
    group_by(!!rlang::ensym(geog), year) %>% 
    summarize(sales_frq = n())
  
  #transaction count with outliers removed
  trans_filtered = df |> 
    mutate(sales2_id = case_when(restype == "CON" & price >= 50000 & price <= 15000000 ~ 1,
                                restype != "CON" & price >= 50000 & price <= 100000000 ~ 1)) |>
    filter(sales2_id == 1) |> 
    group_by(!!rlang::ensym(geog), year) %>% 
    #calculating total sales and median price for sales excluding price outliers
    summarize(sales_frq2 = n(),
              med_pr = sum(price),
              med_pr_adj = sum(price_adj)) |> 
    mutate(med_pr = ifelse(sales_frq2 < 5, NA, med_pr),
           med_pr_adj = ifelse(sales_frq2 < 5, NA, med_pr_adj))
  
  
  data_summary = df |> 
    #creating flags to calculate variables
    mutate(
      #all investors
      inv_trans_f = ifelse(investor_type_purchase != 'Non-investor', 1, 0),
      #investors by size (within 5 year window)
      sm_inv_f = ifelse((year >= 2004 & year <= 2021) & investor_type_purchase == 'Small', 1, 0),
      med_inv_f = ifelse((year >= 2004 & year <= 2021) & investor_type_purchase == 'Medium', 1, 0),
      lrg_inv_f = ifelse((year >= 2004 & year <= 2021) & investor_type_purchase == 'Large', 1, 0),
      inst_inv_f = ifelse((year >= 2004 & year <= 2021) & investor_type_purchase == 'Institutional', 1, 0),
      #investors by type
      inv_llc_f = ifelse(investor_type_purchase_llc == 'Small LLC', 1, 0),
      v_inv_f = ifelse(investor_type_purchase_value != 'Non-value investor', 1, 0),
      c_inv_f = ifelse((year >= 2004 & year <= 2021) & investor_type_purchase_count != 'Non-count investor', 1, 0),
      b_inv_f = ifelse(investor_type_purchase_count != 'Non-count investor', 1, 0),
      #flips within 2 year flip window
      flip_f = ifelse((year >= 2002 & year <= 2023) & flip_ind == 1, 1, 0),
      #condo purchases within 5 year investor window
      con_f = ifelse((year >= 2004 & year <= 2021) & restype == 'CON', 1, 0),
      #condo purchases in 5 year window by investors
      con_inv_f = ifelse((year >= 2004 & year <= 2021) & restype == 'CON' & investor_type_purchase != 'Non-investor', 1, 0),
      #single family purchases within 5-year investor window
      sf_f = ifelse((year >= 2004 & year <= 2021) & restype == 'R1F', 1, 0),
      #single family purchases by investors within 5-year window
      sf_inv_f = ifelse((year >= 2004 & year <= 2021) & restype == 'R1F' & investor_type_purchase != 'Non-investor', 1, 0),
      #two-family purchases in 5 year window
      r2f_f = ifelse((year >= 2004 & year <= 2021) & restype == 'R2F', 1, 0),
      #two-family purchases by investors in 5 year window
      r2f_inv_f = ifelse((year >= 2004 & year <= 2021) & restype == 'R2F' & investor_type_purchase != 'Non-investor', 1, 0),
      #three-family purchases within 5 year window
      r3f_f = ifelse((year >= 2004 & year <= 2021) & restype == 'R3F', 1, 0),
      #three-family purchases by investors withing 5 year window
      r3f_inv_f = ifelse((year >= 2004 & year <= 2021) & restype == 'R3F' & investor_type_purchase != 'Non-investor', 1, 0),
      #investor purchases within 5-year window that were not flips
      inv_non_flip_f = ifelse((year >= 2004 & year <= 2021) & investor_type_purchase != 'Non-investor' & flip_ind == 0, 1, 0),
      #investor purchases within 5-year window that were flips
      inv_flip_f = ifelse((year >= 2004 & year <= 2021) & investor_type_purchase != 'Non-investor' & flip_ind == 1, 1, 0)
    ) |> 
    group_by(!!rlang::ensym(geog), year) |>  
    summarize(
      #count of purchases that were investor transactions
      inv_trans = sum(inv_trans_f),
      #count of purchases that were investor transactions by size
      sm_inv = sum(sm_inv_f),
      med_inv = sum(med_inv_f),
      lrg_inv = sum(lrg_inv_f),
      inst_inv = sum(inst_inv_f),
      #count of purchases that were investor transactions by type
      llc_inv = sum(inv_llc_f),
      value_inv = sum(v_inv_f),
      count_inv = sum(c_inv_f),
      build_inv = sum(b_inv_f),
      #count of purchases that were cash sales
      cash_trans = sum(cash_sale),
      #count of purchases that were flips
      flip_count = sum(flip_f),
      #count of purchases by investors that were not flips
      inv_non_flip = sum(inv_non_flip_f),
      #count of purchases by non-investors that were flips
      inv_flip = sum(inv_flip_f),
      #count of purchases of condos
      con_trans = sum(con_f),
      #count of condo purchases by investors
      con_inv = sum(con_inv_f),
      #count of purchases of single family homes
      sf_trans = sum(sf_f),
      #count of single family home purchases by investors
      sf_inv = sum(sf_inv_f),
      #count of purchases of two family homes
      r2f_trans = sum(r2f_f),
      #count of two family home purchases by investors
      r2f_inv = sum(r2f_inv_f),
      #count of purchases of 3 family homes
      r3f_trans = sum(r3f_f),
      #count of three family home purchases by investors
      r3f_inv = sum(r3f_inv_f)
      )


#join summary tables
warren_join = full_join(warren_all, transactions_all, by = c({{geog}}, 'year')) |> 
    full_join(trans_filtered, by = c({{geog}}, 'year')) |> 
    full_join(data_summary, by = c({{geog}}, 'year')) |> 
    #calculating %s
    mutate(inv_p = inv_trans/sales_frq, #inv trans/tot trans
           sm_inv_p = sm_inv/inv_trans, #small inv trans/tot inv trans
           med_inv_p = med_inv/inv_trans, #med inv trans/tot inv trans
           lrg_inv_p = lrg_inv/inv_trans, #large inv trans/tot inv trans
           inst_inv_p = inst_inv/inv_trans, #inst inv trans/tot inv trans
           li_inv = lrg_inv + inst_inv, #tot large and inst trans
           li_inv_p = li_inv/inv_trans, #lrg + inst inv trans/tot inv trans
           llc_inv_p = llc_inv/inv_trans, #llc inv trans/tot inv trans
           val_inv_p = value_inv/inv_trans, #value inv trans/tot inv trans
           c_inv_p = count_inv/inv_trans, #count investor transactions/tot inv trans
           bld_inv_p = build_inv/inv_trans, #building inv trans/tot inv trans
           cash_p = cash_trans/sales_frq, #cash trans / tot trans
           flip_p = flip_count/sales_frq, #flip trans / tot trans
           con_inv_p = con_inv/con_trans, #inv condo trans/total condo trans
           sf_inv_p = sf_inv/sf_trans, #inv sf trans/total sf trans
           r2f_inv_p = r2f_inv/r2f_trans, #inv 2 family trans/total 2 family trans
           r3f_inv_p = r3f_inv/r3f_trans, #inv 3 family trans/total 3 family trans
    ) %>%
      relocate("li_inv", .after = inst_inv) %>%
      relocate(c("flip_count", "con_inv", "sf_inv", "r2f_inv", "r3f_inv"), .after = cash_trans) %>%
      relocate(c("flip_p", "con_inv_p", "sf_inv_p", "r2f_inv_p", "r3f_inv_p"), .after = cash_p) |> 
      #overwriting data for geographies with less then 5 sales
      mutate(
        med_pr = ifelse(sales_frq2 < 5, NA, med_pr),
        med_pr_adj = ifelse(sales_frq2 < 5, NA, med_pr)
      )

return(warren_join)
}

```


```{r, message=FALSE}
setwd(export_path)

#run ct tables
annual_ct20_agg <- create_tables(warren_select, 'ct20_id')

ct20_publish <- annual_ct20_agg |> 
   arrange(ct20_id, year)

#export ct table
write.csv(ct20_publish, 'ct20_annual_2000_2025.csv')

annual_ct10_agg <- create_tables(warren_select, 'ct10_id')

ct10_publish <- annual_ct10_agg |> 
   arrange(ct10_id, year)

#export ct table
write.csv(ct10_publish, 'ct10_annual_2000_2025.csv')
```

