---
title: "WGBH"
output: html_document
---

To do: re-run and check 

Request:
1. the number of purchases and the percentages of transactions by those investors
- for the entire study period and 2015-2025
2. the names of investors and their businesses and addresses for properties purchased
- LLC names and addresses; time period: 2015-2025 

Main geography: Lynn
Also interested: Haverhill, Lowell

```{r, warning=FALSE}
library(tidyverse)
options('scipen' = 10)

data_path = "K:/DataServices/Projects/Current_Projects/Regional_Plan_Update_Research/Speculative Investment/Data/"
export_path = "K:/DataServices/Projects/Data_Requests/2026/WGBH_Neisloss"

setwd(data_path)

#starting from mapc filtered file
warren <- read_csv("2000_2025_warren_speculative-investment-analysis-dataset_withforeclosure_5yr-window-networks.csv")

warren_clean <- warren |> 
  #clean up issue of missing inv_type_purchase due to NAs
  mutate(
    #make NAs a string
    investor_type_purchase_value = ifelse(is.na(investor_type_purchase_value), "NA", investor_type_purchase_value),
    #redefine final investor category for cases where is is missing
    investor_type_purchase = ifelse(!is.na(investor_type_purchase), investor_type_purchase, 
                              ifelse((investor_type_purchase_value=='Institutional' | investor_type_purchase_count=='Institutional'),'Institutional',
                                ifelse((investor_type_purchase_value=='Large' | investor_type_purchase_count=='Large'), 'Large',
                                  ifelse((investor_type_purchase_value=='Medium'| investor_type_purchase_count=='Medium'), 'Medium',
                                   ifelse((investor_type_purchase_value=='Small'| investor_type_purchase_count=='Small' | investor_type_purchase_llc=='Small LLC' | investor_type_purchase_building=='Building Investor'),'Small', 'Non-investor')))))
  )

rm(warren)
gc()
```

#Identify investors
```{r}
investors <- warren_clean |> 
  #filter all transactions to given year and munis
  filter(year >= 2015) |>
  #select needed variables
  select(buyer1_adj, municipal, investor_type_purchase, investor_type_purchase_llc, investor_type_purchase_building, investor_type_purchase_count, investor_type_purchase_value, network_group) |> 
  #filter to only LLC investors
  filter(investor_type_purchase_llc == "Small LLC")

lynn_investors <- investors |> 
  filter(municipal == "Lynn") |> 
  #remove duplicates
  distinct() 
  
haverhill_investors <- investors |> 
  filter(municipal == "Haverhill") |> 
  #remove duplicates
  distinct() 
  
lowell_investors <- investors |> 
  filter(municipal == "Lowell") |> 
  #remove duplicates
  distinct() 
  
all_investors <- rbind(lynn_investors, haverhill_investors, lowell_investors)

#how many ownership networks do these investors really represent?
investors |> count(network_group)

setwd(export_path)
write.csv(lynn_investors, "lynn_investors.csv")
write.csv(haverhill_investors, "haverhill_investors.csv")
write.csv(lowell_investors, "lowell_investors.csv")

rm(investors, lynn_investors, haverhill_investors, lowell_investors)
gc()
```
1,654 unique LLC buyer names across 543 network groups
"all_investors" will include duplicates when investors have made purchases in more than one of the 3 included municipalities

#Network table filtered to investors of interest - not actually using this
```{r}
# Get crosswalk from network to owner
who_owns_path <- "C:/Users/aderosa/OneDrive - Metropolitan Area Planning Council/Shared Documents - Data Services/_Current Projects/_Housing/LLC Owner Networks - NNIP/04_Data/who-owns-csvs/"
setwd(who_owns_path)
owners_networks <- read_csv("owners.csv")

owners_inst <- owners_networks |>
  filter(inst == TRUE) |>
  distinct(name, cosine_group, network_group)

rm(owners_networks)
# what information is in other network analysis tables that would be of interest for this request?
```

#Statewide purchases by Lynn, Haverhill, Lowell investors
```{r}
#total number of purchases made by the investor in MA
ma_inv_purchases <- warren_clean |> 
  select(buyer1_adj, network_group, year) |> 
  filter(buyer1_adj %in% all_investors$buyer1_adj) |> 
  group_by(buyer1_adj) |> 
  #count for full study period
  mutate(ma_inv_pur_00_25 = n()) |> 
  filter(year >= 2015) |> 
  #count 2015-2025
  mutate(ma_inv_pur_15_25 = n()) |> 
  select(-c(year)) |> 
  distinct()

#total number of purchases made by the investor network in MA
ma_ntwk_purchases <- warren_clean |> 
  select(network_group, year, investor_type_purchase) |> 
  filter(network_group %in% all_investors$network_group & !is.na(network_group)) |> 
  group_by(network_group) |> 
  #count for full study period
  mutate(ma_inv_pur_00_25 = n()) |> 
  filter(year >= 2015) |> 
  #count 2015-2025
  mutate(ma_inv_pur_15_25 = n()) |> 
  select(-c(year)) |> 
  distinct()

setwd(export_path)
write.csv(ma_inv_purchases, "ma_investor_purchase_summary.csv")
write.csv(ma_ntwk_purchases, "ma_network_purchase_summary.csv")
gc()
```
1,654 uniquely named investor llcs, 543 unique networks
#municipal purchases by investors
```{r}
warren_muni <- warren_clean |> 
  filter(municipal == "Lynn" | municipal == "Haverhill" | municipal == "Lowell") 

#count of purchases by investor by muni
inv_purchases <- warren_muni |> 
  select(buyer1_adj, network_group, municipal, year) |> 
  group_by(municipal) |> 
  #muni transaction count for full study period
  mutate(muni_00_25 = n(),
         inv_f_00_25 = ifelse(buyer1_adj %in% all_investors$buyer1_adj, 1, 0)) |> 
  filter(year >= 2015) |> 
  #muni transaction count 2015-2025
  mutate(muni_15_25 = n(),
         inv_f_15_25 = ifelse(buyer1_adj %in% all_investors$buyer1_adj, 1, 0)) |> 
  select(-c(year)) |> 
  group_by(municipal, buyer1_adj) |> 
  mutate(
    #count inv purchases in muni
    inv_pur_00_25 = sum(inv_f_00_25),
    #percent inv purchases in muni
    p_inv_00_25 = sum(inv_f_00_25)/muni_00_25,
    #count inv purchases in muni (2015-2025)
    inv_pur_15_25 = sum(inv_f_15_25),
    #percent inv purchases in muni (2015-2025)
    p_inv_15_25 = sum(inv_f_15_25)/muni_15_25,
  ) |> 
  select(-c(inv_f_00_25, inv_f_15_25))

#count of purchases by investor by muni
ntwk_purchases <- warren_muni |> 
  select(network_group, municipal, year) |> 
  group_by(municipal) |> 
  #muni transaction count for full study period
  mutate(muni_00_25 = n(),
         inv_f_00_25 = ifelse(network_group %in% all_investors$network_group, 1, 0)) |> 
  filter(year >= 2015) |> 
  #muni transaction count 2015-2025
  mutate(muni_15_25 = n(),
         inv_f_15_25 = ifelse(network_group %in% all_investors$network_group, 1, 0)) |> 
  select(-c(year)) |> 
  group_by(municipal, network_group) |> 
  mutate(
    #count inv purchases in muni
    inv_pur_00_25 = sum(inv_f_00_25),
    #percent inv purchases in muni
    p_inv_00_25 = sum(inv_f_00_25)/muni_00_25,
    #count inv purchases in muni (2015-2025)
    inv_pur_15_25 = sum(inv_f_15_25),
    #percent inv purchases in muni (2015-2025)
    p_inv_15_25 = sum(inv_f_15_25)/muni_15_25,
  ) |> 
  select(-c(inv_f_00_25, inv_f_15_25))

setwd(export_path)
write.csv(inv_purchases, "investor_purchase_summary.csv")
write.csv(ntwk_purchases, "network_purchase_summary.csv")
gc()
```

